Formal Synthesis of Digital Controllers for Observable Continuous Plants with Transient Performance Specifications

### 2018-02-01 CAV2018/paper.tex
submitted to CAV 2018
proper macro for log

\begin{theorem}
%
Given a set of eigenvalues for a discrete time model of a $p^{th}$ order
stable system $\{\lambda_i =e^{\omega_i} \mid i \in \{1, \ldots,p\},
|\lambda_i| \leq 1 \}$, the settling time of the system is upper-bounded by
%
\begin{equation}
%
t_s<-log_{\overline{\lambda}}({P_s}), \text{ where }
  \overline{\lambda} = \max(|\lambda_i|) \text{ and }
  \overline{\lambda}<1,\ \forall i \in \{1, \ldots,p\}
%
\label{eq:set_time}
\end{equation}
%
where $0\leq P_s \leq 1$ is the desired portion of the target within which
the output must remain after $t_s$.  The proof is in
Appendix~\ref{sec:proofs}.
%
\end{theorem}

@@

\begin{table}
\centering
\footnotesize
%
\begin{tabular}{| r | l | c | c | r | r | r | r | r | r | r | r | r |}
%
\hline
\# & \multicolumn{1}{|c|}{Benchmark} & \multicolumn{1}{|c|}{order} & \multicolumn{1}{|c|}{No.} & \multicolumn{4}{|c|}{CEGIS} & \multicolumn{4}{|c|}{CEGIS-OR} \\
   &                                 & & \multicolumn{1}{|c|}{}  & \multicolumn{1}{|c|}{Min} & \multicolumn{1}{|c|}{Median} & \multicolumn{1}{|c|}{t.o.} & \multicolumn{1}{|c|}{found} & \multicolumn{1}{|c|}{Min} & \multicolumn{1}{|c|}{Median} & \multicolumn{1}{|c|}{t.o.} & \multicolumn{1}{|c|}{found}\\\hline
1  & DC Motor          & 2 & 100 & 1.3\,s & 26.9\,s& 22& 20& 1.1\,s &   3.8\,s & 2 & 58\\
2  & Helicopter        & 3  & 100 & 57.2\,s & t.o. & 54 & 15 & 7.1\,s  &   18.0\,s & 9 & 90\\
3  & Inverted Pendulum & 4 & 100 &  277\,s& t.o. & 59& 0 & 20.0\,s  &  t.o. & 50 & 20\\
4  & Magnetic Pointer  & 2  & 100 & 5.7\,s & 229.9\,s & 38 & 42 & 3.0\,s  &  9.5\,s & 5 & 95 \\
%5 & Magnetic Suspension          & 2 & 1 & 12.2\,s& 0.6\,s & & & N/A  &   N/A &   N/A\\
5  & Pendulum          & 2 & 100 & 0.8\,s & 53.6\,s & 27 & 71 &0.4\,s  &   4.9\,s & 6 & 92\\
6  & Suspension        & 4 & 100 &15.0\,s & 177.2\,s & 30 & 70 & 14.8\,s&166.9\,s  & 26 & 73\\
7  & Tape Driver       & 3 & 100 & 4.0\,s& 17.7\,s & 11 & 80 &3.1\,s  &   16.8\,s & 8 & 82\\
%9 & Satellite          & 2 & 1 & 9.4\,s& 0.7\,s & & & N/A  &   N/A &   N/A\\
%   & Total                 &   & 2002 & & & & & 1.2\,s & 16.1\,s & 876.8\,s\\
\hline
%
\end{tabular}
\vspace{0.05in}
\caption{\label{tab:cegis_results}
Experimental results.  Minimum, and Median times for standard CEGIS vs
CEGIS-OR on multiple benchmarks.  No.  indicates Number of benchmarks,
t.o.~indicates timeouts (in the median case, it indicates more than half of
the benchmarks timed out).  The order denotes the number of variables of a
model.}
%
\end{table}

@@

%-------------------------------
\subsubsection{Transient response} 
\label{sec:transientspecificationproof}
%-------------------------------
\begin{theorem}
Given a set of eigenvalues for a discrete time model of an $p^{th}$ order system $\lambda_i =e^{\omega_i}, i \in [1, ...\,,p]$, the settling time of the system is upper-bounded by
\begin{equation}
t_s<-log_{\overline{\lambda}}({P_s}), \text{ where }  \overline{\lambda} = \max(|\lambda_i|) \text{ and } \overline{\lambda}<1,\ \forall i \in \{1, \ldots,p\}
\label{eq:set_time}
\end{equation}
where $0\leq P_s \leq 1$ is the desired percentage of the target within which the output must remain after $t_s$.
\end{theorem}
\begin{proof}
In a SISO system, there are two values that typically define the
response-time of the system (using a step response).
%
\begin{enumerate}
%
\item The \emph{rise time} is defined as the time it takes the output to
reach a percentage (e.g., $90\%$) of the target value given a step change in
the input.
%
\item The \emph{settling time} is defined as the time it takes the output to
stabilize within a margin of the target value (e.g., $10\%$).
%
\end{enumerate} 
%
%In the case of first and second order systems, the decay is governed by the equations 
%%
%\begin{align*}
%&y_{decay}=
%&\left\{
%\begin{array}{cc}
%e^{-|\omega| t}& 1^{st} order\\
%e^{\zeta |\omega| t}\left(\frac{\sin\left(|\omega| t\sqrt{1-\zeta^2}+\sin^{-1}\sqrt{1-\zeta^2}\right)}{\sqrt{1-\zeta^2}} \right)&\zeta<1\\
%e^{-|\omega| t}-|\omega| t e^{-|\omega| t}&\zeta=1\\
%\frac{|\omega| e^{-|\omega_2| t}-|\omega_2| e^{-|\omega| t}}{|\omega|-|\omega_2|}&\zeta>1\end{array}\right.
%\end{align*}
%where $|\omega|$ and $|\omega_2|$ are the magnitudes of the eigenvalues (roots) of the system and  $\zeta=-\cos(\theta)$ where $\theta$ is the polar angle of the conjugate pair if it exists ($\zeta \geq 1$ if it doesn't).
%These equations become increasingly complicated for higher order systems. However, we note their strong dependency on the magnitude of the eigenvalues, and use this property to define an upperbound on the decay function.
%All of the above equations will decay faster than $e^{-|\omega| t}$ given $|\omega| < |\omega_2|$, hence we may use this function as our upper bound.

@@

In the case where we allow for geometric multiplicities, we need to
compensate for the effects of $\binom{k}{j}$ \footnote{Tighter bounds can be
obtained by evaluating $\frac{\sum\limits_{j=1}^{gm(i)}c_{i+j}\binom{k}{j}
\lambda_i^{-j}{\vec{x}_0}_j}{\sum\limits_{j=1}^{gm(i)}c_{i+j}{\vec{x}_0}_j}$.}. 
We recall here that our interest is to reach within $10\%$ of the final
value, and that for a given $k>\underline{k}$ the term $\lambda^k$ will
converge faster than $\binom{k}{j}$ diverges.  The requirements for
$|\overline{\omega}|$ are calculated as follows:
%
\begin{itemize}
%
\item Find $\underline{k}=\frac{t_s}{T_s}$ where $t_s$ is the settling time and $T_s$ the sampling time.
%
\item Calculate $|\overline{\omega}|=log_{t_s}\left(P_s\binom{\underline{k}}{gm-1}^{-1}\right)$,
where $gm$ is the maximum desired geometric multiplicity.
%
\item Ensure that $|\overline{\omega}| \leq \frac{\underline{k}}{\underline{k}+1}$ and that the
synthesized closed loop does not have a larger $gm$ than specified.
%
\end{itemize}
%\qed
\end{proof}

>>>>

\begin{theorem}
%
Given a set of eigenvalues for a discrete time model of a $p^{th}$ order
stable system $\{\lambda_i =e^{\omega_i} \mid i \in \{1, \ldots,p\},
|\lambda_i| \leq 1 \}$, the settling time of the system is upper-bounded by
%
\begin{equation}
%
t_s<-\log_{\overline{\lambda}}({P_s}), \text{ where }
  \overline{\lambda} = \max(|\lambda_i|) \text{ and }
  \overline{\lambda}<1,\ \forall i \in \{1, \ldots,p\}
%
\label{eq:set_time}
\end{equation}
%
where $0\leq P_s \leq 1$ is the desired portion of the target within which
the output must remain after $t_s$.  The proof is in
Appendix~\ref{sec:proofs}.
%
\end{theorem}

@@

\begin{table}
\centering
\footnotesize
%
\begin{tabular}{| r | l | c | c | r | r | r | r | r | r | r | r | r |}
%
\hline
\# & \multicolumn{1}{|c|}{Benchmark} & \multicolumn{1}{|c|}{order} & \multicolumn{1}{|c|}{No.} & \multicolumn{4}{|c|}{CEGIS} & \multicolumn{4}{|c|}{CEGIS-OR} \\
   &                                 & & \multicolumn{1}{|c|}{}  & \multicolumn{1}{|c|}{Min} & \multicolumn{1}{|c|}{Median} & \multicolumn{1}{|c|}{t.o.} & \multicolumn{1}{|c|}{found} & \multicolumn{1}{|c|}{Min} & \multicolumn{1}{|c|}{Median} & \multicolumn{1}{|c|}{t.o.} & \multicolumn{1}{|c|}{found}\\\hline
1  & DC Motor          & 2 & 100 & 1.3\,s & 26.9\,s& 22& 20& 1.1\,s &   3.8\,s & 2 & 58\\
2  & Helicopter        & 3  & 100 & 57.2\,s & t.o. & 54 & 15 & 7.1\,s  &   18.0\,s & 9 & 90\\
3  & Inverted Pendulum & 4 & 100 &  277\,s& t.o. & 82& 0 & 20.0\,s  &  t.o. & 55 & 20\\
4  & Magnetic Pointer  & 2  & 100 & 5.7\,s & 229.9\,s & 38 & 42 & 3.0\,s  &  9.5\,s & 5 & 95 \\
%5 & Magnetic Suspension          & 2 & 1 & 12.2\,s& 0.6\,s & & & N/A  &   N/A &   N/A\\
5  & Pendulum          & 2 & 100 & 0.8\,s & 53.6\,s & 27 & 71 &0.4\,s  &   4.9\,s & 6 & 92\\
6  & Suspension        & 4 & 100 &15.0\,s & 177.2\,s & 30 & 70 & 14.8\,s&166.9\,s  & 26 & 73\\
7  & Tape Driver       & 3 & 100 & 4.0\,s& 17.7\,s & 11 & 80 &3.1\,s  &   16.8\,s & 8 & 82\\
%9 & Satellite          & 2 & 1 & 9.4\,s& 0.7\,s & & & N/A  &   N/A &   N/A\\
%   & Total                 &   & 2002 & & & & & 1.2\,s & 16.1\,s & 876.8\,s\\
\hline
%
\end{tabular}
\vspace{0.05in}
\caption{\label{tab:cegis_results}
Experimental results.  Minimum, and Median times for standard CEGIS vs
CEGIS-OR on multiple benchmarks.  No.  indicates Number of benchmarks,
t.o.~indicates timeouts (in the median case, it indicates more than half of
the benchmarks timed out).  The order denotes the number of variables of a
model.}
%
\end{table}

@@

%-------------------------------
\subsubsection{Transient response} 
\label{sec:transientspecificationproof}
%-------------------------------
\begin{theorem}
Given a set of eigenvalues for a discrete time model of an $p^{th}$ order system $\lambda_i =e^{\omega_i}, i \in [1, ...\,,p]$, the settling time of the system is upper-bounded by
\begin{equation}
t_s<-\log_{\overline{\lambda}}({P_s}), \text{ where }  \overline{\lambda} = \max(|\lambda_i|) \text{ and } \overline{\lambda}<1,\ \forall i \in \{1, \ldots,p\}
\label{eq:set_time}
\end{equation}
where $0\leq P_s \leq 1$ is the desired percentage of the target within which the output must remain after $t_s$.
\end{theorem}
\begin{proof}
In a SISO system, there are two values that typically define the
response-time of the system (using a step response).
%
\begin{enumerate}
%
\item The \emph{rise time} is defined as the time it takes the output to
reach a percentage (e.g., $90\%$) of the target value given a step change in
the input.
%
\item The \emph{settling time} is defined as the time it takes the output to
stabilize within a margin of the target value (e.g., $10\%$).
%
\end{enumerate} 
%
%In the case of first and second order systems, the decay is governed by the equations 
%%
%\begin{align*}
%&y_{decay}=
%&\left\{
%\begin{array}{cc}
%e^{-|\omega| t}& 1^{st} order\\
%e^{\zeta |\omega| t}\left(\frac{\sin\left(|\omega| t\sqrt{1-\zeta^2}+\sin^{-1}\sqrt{1-\zeta^2}\right)}{\sqrt{1-\zeta^2}} \right)&\zeta<1\\
%e^{-|\omega| t}-|\omega| t e^{-|\omega| t}&\zeta=1\\
%\frac{|\omega| e^{-|\omega_2| t}-|\omega_2| e^{-|\omega| t}}{|\omega|-|\omega_2|}&\zeta>1\end{array}\right.
%\end{align*}
%where $|\omega|$ and $|\omega_2|$ are the magnitudes of the eigenvalues (roots) of the system and  $\zeta=-\cos(\theta)$ where $\theta$ is the polar angle of the conjugate pair if it exists ($\zeta \geq 1$ if it doesn't).
%These equations become increasingly complicated for higher order systems. However, we note their strong dependency on the magnitude of the eigenvalues, and use this property to define an upperbound on the decay function.
%All of the above equations will decay faster than $e^{-|\omega| t}$ given $|\omega| < |\omega_2|$, hence we may use this function as our upper bound.

@@

In the case where we allow for geometric multiplicities, we need to
compensate for the effects of $\binom{k}{j}$ \footnote{Tighter bounds can be
obtained by evaluating $\frac{\sum\limits_{j=1}^{gm(i)}c_{i+j}\binom{k}{j}
\lambda_i^{-j}{\vec{x}_0}_j}{\sum\limits_{j=1}^{gm(i)}c_{i+j}{\vec{x}_0}_j}$.}. 
We recall here that our interest is to reach within $10\%$ of the final
value, and that for a given $k>\underline{k}$ the term $\lambda^k$ will
converge faster than $\binom{k}{j}$ diverges.  The requirements for
$|\overline{\omega}|$ are calculated as follows:
%
\begin{itemize}
%
\item Find $\underline{k}=\frac{t_s}{T_s}$ where $t_s$ is the settling time and $T_s$ the sampling time.
%
\item Calculate $|\overline{\omega}|=\log_{t_s}\left(P_s\binom{\underline{k}}{gm-1}^{-1}\right)$,
where $gm$ is the maximum desired geometric multiplicity.
%
\item Ensure that $|\overline{\omega}| \leq \frac{\underline{k}}{\underline{k}+1}$ and that the
synthesized closed loop does not have a larger $gm$ than specified.
%
\end{itemize}
%\qed
\end{proof}
